---
title: "trail_mlr"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# Loading Libraries

```{r}
library(varhandle)

```

# Loading Data

```{r}
# reads dataset
data <- read.csv2("../data/cardio.csv")

# drops 'id' attribute from the dataset
data <- subset(data, select = -c(id))

# converts 'weight' from character to numeric
data$weight <- unfactor(data$weight)

# converts unit of 'age' from days to years
data$age <- round(data$age / 365, digits = 0)

# creates 'BMI' column from corresponding 'height' and 'weight'
data$BMI <- (data$weight / ((data$height * 0.01) ^ 2))
#round BMI
data$BMI <- round(data$BMI, digits = 1)

# changes column names to readable format
colnames(data) <- c("Age", "Gender", "Height", "Weight", "Systole", "Diastole", "Cholesterol", "Glucose", "Smoking", "Alcohol", "Active", "Target", "BMI")
head(data)
```

# Exploratory Analysis


# Managing Data

```{r}
# removes observation for which 'Systole' or 'Diastole' has negative values
data <- subset(data, Systole > 0 & Diastole > 0)

# removes 'ap_hi' and 'ap_lo' outliers
iqr_multiplier = 3
ap_hi_upper_threshold <- quantile(data$Systole, .75) + iqr_multiplier * (quantile(data$Systole, .75) - quantile(data$Systole, .25))
ap_hi_lower_threshold <- quantile(data$Systole, .25) - iqr_multiplier * (quantile(data$Systole, .75) - quantile(data$Systole, .25))
ap_lo_upper_threshold <- quantile(data$Diastole, .75) + iqr_multiplier * (quantile(data$Diastole, .75) - quantile(data$Diastole, .25))
ap_lo_lower_threshold <- quantile(data$Diastole, .25) - iqr_multiplier * (quantile(data$Diastole, .75) - quantile(data$Diastole, .25))
data <- subset(data, Systole < ap_hi_upper_threshold & Systole > ap_hi_lower_threshold & Diastole < ap_lo_upper_threshold & Diastole > ap_lo_lower_threshold)

# removes 'weight' outliers
weight_iqr_multiplier = 1.5
weight_upper_threshold <- quantile(data$Weight, .75) + weight_iqr_multiplier * (quantile(data$Weight, .75) - quantile(data$Weight, .25))
weight_lower_threshold <- quantile(data$Weight, .25) - weight_iqr_multiplier * (quantile(data$Weight, .75) - quantile(data$Weight, .25))
data <- subset(data, Weight > weight_lower_threshold & Weight < weight_upper_threshold)

# summerizes newly encoded dataset
summary(data)
```

# Modeling Data

## change blood pressure into three levels:

```{r}
###    lower than normal:-1, normal:0, higher than normal: 1
data$BloodPressure =ifelse((data$Systole>=140 | data$Diastole>=90),1,ifelse((data$Systole<=90 | data$Diastole<=60),-1,0))
```

## multiple linear regression on factors of target
```{r}
library(dplyr)
sample1= sample_n(data,1000)
mlr =lm(data = sample1,Target~Age+Gender+BMI+Cholesterol+Glucose+Smoking+Alcohol+Active+BloodPressure)
summary(mlr)
```

## factor selection according to parsimony principle
```{r}
###find collinearity among factors
library(GGally)
ggpairs(data = sample1,c(1:2,7:11,13,14))
```
abvious collinearity: glucose~cholesterol (0.518), Smoking~Gender(0.32), Alcohol~Gender(0.206), Alcohol~Smoking(0.346) , BloodPressure~BMI (0.215)

### also remove the following factors one by one whose t value is lower than 1.5 (and even the absolute values of standard errors are all larger than Estimate!) and thus significance is really low:
### alcohol, smoking, gender
```{r}

mlr2 =lm(data = sample1,Target~Age+BMI+Cholesterol+Glucose+Active+BloodPressure)
summary(mlr2)
```

## After trying removing collineared factors BloodPressure or BMI which resulted in less adjusted R-squared, we could tentatively conlude that the following factors remain influencial:

## Age + BMI + Glucose + Active + BloodPressure

_____________________

